Solving Consensus with OpenFlow

### 2015-02-17 inband-STN.tex
stefan pass

\documentclass[11pt,pdftex,letter]{article}
%\documentclass[11pt]{llncs}
%\documentclass[11pt,pdftex]{article}
%\documentclass{sig-alternate-10pt}
%\usepackage{amsthm}
%\usepackage{algorithm}
%\usepackage[noend]{algorithmic}
\usepackage[lined,boxed,commentsnumbered]{algorithm2e}

\usepackage{amssymb}
\usepackage{comment}
\usepackage{amsmath}
%\usepackage{graphicx}
%\usepackage{color}
\usepackage{fullpage}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsRule{*}{mps}{*}{<++>}

\ifx\pdftexversion\undefined
\usepackage[dvips]{graphicx}
\else
  \usepackage[pdftex]{graphicx}
  \DeclareGraphicsRule{*}{mps}{*}{}
\fi

\def\lf{\tiny}
\def\rrnnll{\setcounter{linenumber}{0}}
\def\nnll{\refstepcounter{linenumber}\lf\thelinenumber}
\newcounter{linenumber}

\usepackage[usenames,dvipsnames]{color}
%\usepackage[caption=true,font=footnotesize]{subfig}
\usepackage{subfigure}
%\usepackage{xspace} 	% Guesses whether a space is needed when invoked
\usepackage{listings}
\usepackage{url}
\usepackage{wrapfig}
\usepackage{cite}
%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
%
\usepackage[boxed]{algorithm2e}
%\usepackage{comment}

@@

\newcommand{\CPO}{\textsc{FixTag}}
\newcommand{\DPO}{\textsc{ReuseTag}}
\newcommand{\PS}{\textsc{PS}}
\newcommand{\Bit}{\textsc{Bit}}

\newcommand{\gns}{Global Network State\xspace}
\newcommand{\GNS}{\textsc{GNS}\xspace}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{lemma}{Lemma}[section]
%\newtheorem{claim}{Claim}[section]
%\newtheorem{scenarios}{Scenarios}[section]
%\newtheorem{observation}{Observation}
%\newtheorem{takeaway}{Takeaway}[section]
%\newtheorem{definition}{Definition}

%
%
% carry over Herald's group cool way of marking changes
\definecolor{heraldBlue}{rgb}{0.0,0.0,0.8}
\definecolor{heraldRed}{rgb}{0.8,0.0,0.0}
\definecolor{heraldGray}{rgb}{0.4,0.4,0.4}
\definecolor{heraldBlack}{rgb}{0.0,0.0,0.0} %removes comment color
\definecolor{heraldGreen}{rgb}{0.0,0.4,0.0} %removes comment color
\def\r#1{\textcolor{heraldBlue}{\em #1}}
\def\q#1{\textcolor{heraldRed}{\em #1}}
\def\d#1{\textcolor{heraldBlue}{#1}}
\def\R#1{\textcolor{heraldBlue}{#1}}
\def\D#1{\textcolor{heraldBlue}{#1}}
%
%\DeclareMathOperator{\respc}{res}

\newcommand{\len}{\text{d}}

\newcommand{\nodes}{\mathcal{N}}
\newcommand{\links}{\mathcal{E}}
\newcommand{\epoints}{\Pi}
\newcommand{\epoint}{\pi}
\newcommand{\legsw}{\mathcal{L}}
\newcommand{\sdnsw}{\mathcal{S}}
\newcommand{\sw}{\legsw\sqcup\sdnsw}
\newcommand{\fset}{FS}

\newcommand{\Cost}{\gamma}

\newcommand{\ft}{ft}

\newcommand{\eepath}{p}
\newcommand{\link}{e}

\newcommand{\dom}{\textit{dom}}
\newcommand{\pr}{\textit{pr}}
\newcommand{\CPOs}{\textit{paths}}
\newcommand{\seq}{\textit{seq}}
\newcommand{\cur}{\textit{cur}}
\newcommand{\Tag}{\textit{tag}}

\newcommand{\ports}{\Pi}
\newcommand{\seport}{\pi}
\newcommand{\inports}{\overrightarrow{\ports}}
\newcommand{\swports}{\overleftrightarrow{\ports}}
\newcommand{\seinports}{\epoints^{\bullet}}
\newcommand{\seinportsone}{\seinports_1}
\newcommand{\seinportstwo}{\seinports_2}
\newcommand{\seinportsthree}{\seinports_3}
\newcommand{\neinports}{\epoints^{\circ}}

\newcommand{\readt}{\texttt{r}}
\newcommand{\writet}{\texttt{w}}
\newcommand{\op}{\texttt{op}}

\newcommand{\cellblocks}{CB}
\newcommand{\cellblock}{c}
\newcommand{\frontier}{\mathcal{F}}
\newcommand{\MIP}{\textsc{Opt}}
\newcommand{\smartparagraph}[1]{\noindent{\bf #1}\ }
\newcommand{\eg}{{\it e.g.}}
\newcommand{\ie}{{\it i.e.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\etal}{{\it et al.}\xspace}
\newcommand{\id}{{\it id}}

\def\TR{0}
\def\NOTES{1}
\def\SAVESPACE{1}
\def\SHOWAUTHORS{1}
\def\SHOWGIT{0}
% variables may contain the above definitions to control build options at compile-time
%\input{variables}

\if \SAVESPACE 1
%\usepackage{setspace}
%\usepackage{titlesec}
%\titlespacing\section{0pt}{7pt}{6pt}
\usepackage{titlesec}
\titlespacing\section{0pt}{7pt}{6pt}

\fi

\if \NOTES 1
\newcommand{\mcnote}[1]{\textcolor{heraldBlue}{\small \bf [MC: #1]}}
\newcommand{\dlnote}[1]{\textcolor{heraldBlue}{\small \bf [DL: #1]}}
\newcommand{\ssnote}[1]{\textcolor{heraldBlue}{\small \bf [SS: #1]}}
\newcommand{\pknote}[1]{\textcolor{heraldBlue}{\small \bf [PK: #1]}}
%\newcommand{\problem}[1]{\textcolor{heraldRed}{\small \bf [PROBLEM: #1]}}
\else
\newcommand{\mcnote}[1]{}
\newcommand{\dlnote}[1]{}
\newcommand{\ssnote}[1]{}
\newcommand{\pknote}[1]{}
\newcommand{\problem}[1]{}
\fi

%Petr's definitions
\newcommand{\ack}{\textit{ack}}
\newcommand{\nack}{\textit{nack}}
\newcommand{\rmw}{\textit{rmw}}
\newcommand{\State}{\mathit{States}}
\newcommand{\ignore}[1]{}
\newcommand{\Pb}{CPC}
\newcommand{\E}{\mathcal E}
%

%[[PK environments for article style
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
\newenvironment{proofsketch}[1][Proof sketch]{\noindent\textbf{#1.} }{\hfill $\Box$\\[2mm]}
%]]

@@

\author{Liron, Stefan, Petr}

@@

\begin{abstract}
We solve consensus with OpenFlow, using multi-write
\end{abstract}

@@

\section{SDN switch configuration space as shared memory}\label{sec:todo}

@@

\section{The one touch policy update problem}\label{sec:todo}

@@

\section{Dealing with multiple updates}\label{sec:todo}

@@

\bibliographystyle{abbrv}
\bibliography{references}  % main.bib is the name of the Bibliography in this case

>>>>

%\documentclass[11pt,pdftex,letter]{article}
\documentclass[conference]{sigcomm-alternate}

%\documentclass{sig-alternate-10pt}

%\documentclass[11pt]{llncs}
%\documentclass[11pt,pdftex]{article}
%\documentclass{sig-alternate-10pt}
%\usepackage{amsthm}
%\usepackage{algorithm}
%\usepackage[noend]{algorithmic}

\usepackage{amssymb}
\usepackage{comment}
\usepackage{amsmath}
%\usepackage{graphicx}
%\usepackage{color}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsRule{*}{mps}{*}{<++>}

%\usepackage[caption=true,font=footnotesize]{subfig}
%\usepackage{xspace} 	% Guesses whether a space is needed when invoked
\usepackage{cite}
%\usepackage{framed}
\usepackage{framed,color}
\definecolor{shadecolor}{rgb}{0.9,0.9,0.9}
%
%\usepackage{comment}

@@

@@

\title{One-Shot Coordination of Distributed SDN Controllers:\\The Case for Multi-Write Consensus}

\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}

\author{
Liron Schiff\thanks{Supported by the European Research Council (ERC) Starting Grant no. 259085 and by the Israel Science Foundation Grant no.~1386/11.} $^1$, 
Stefan Schmid$^2$, Petr Kuznetsov$^3$ \\
\small $^1$ Tel Aviv University, Israel; $^2$ TU Berlin \& T-Labs, Germany; $^3$ ParisTech, France
}

@@

\begin{abstract}
The control plane of future Software-Defined Networks (SDNs)
is likely to be \emph{distributed}: a redundant control plane
can ensure a high availability, and a spatially distributed
control plane allows to handle time-critical dataplane
events close to their origin. Moreover, future SDNs may also
allow multiple administrators or participatory users to update
policies concurrently, e.g., by offering transactional interfaces. 
However, not much is known today about how to implement 
a consistent distributed control plane which supports concurrent
policy installations.

In this paper, we propose to coordinate distributed controllers
directly on the Openflow switches: exploiting classic results
on multi-writer consensus, we show how to use Openflow 
to synchronize controllers and install policies in a consistent
manner, \emph{in one shot}. Our approach is attractive as it
does not require complicated helping mechanisms in case of controller
failures. 
\end{abstract}

@@

By consolidating and outsourcing the control over the dataplane switches to a logically
centralized controller,
Software-defined networks (SDNs) simplify the network management and facilitate faster
innovations, as the (software) control plane is decoupled and can evolve independently.
Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
by relying on a simple match-action paradigm which allows us to define
forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)
header fields.

While the logically centralized network perspective has the potential to simplify
the network management---many network tasks are inherently non-local---and introduces
many flexibilities,
managing and operating an SDN in a more adaptive but consistent manner is non-trivial, even with SDN.
A particularly interesting problem regards the consistent installation of new policies
or routes.~\cite{todo,todo,todo} While implementing consistent network updates 
is challenging already from the perspective of a single controller, there exists a wide consensus
that the logically centralized SDN control planes will likely be \emph{actually distributed} in the future.

The reason for implementing SDN control planes in a distributed manner are manyfold.
First, in order to provide a high availability as well as a minimal degree of
fault-tolerance, controllers should be redundant~\cite{onix,stn,todo}: a failure
of one controller can be masked by other controllers. Second, it has also been proposed
to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
communication intensive control plane events close to their origin.~\cite{devoflow,kandoo,jukka,disco}
Third, larger SDNs are likely to be operated by multiple administrators or may even offer
participatory interfaces where different users can install and trigger policy changes
concurrently.~\cite{participatory,stn}

However, today, we do not have a good understanding of how to realize
such distributed control planes. The problem is essentially a distributed
computing one: Multiple controllers may simultaneously try to
install conflicting policies, and a consensus must be reached which one of
the two is actually installed. Moreover, controllers may also fail,
and a mechanism must be provided to complete an update in this scenario.

\subsection{Our Contributions}

In this paper, we propose an approach to solve these problems.
In particular, we promote a very direct approach to coordinate controllers,
using the Openflow bundle feature to make an atomic consensus and update:
using one message, a single controller will be able to install its policy,
\emph{in one shot}. Our approach relies on a classic result in 
distributed computing, namely multi-writer consensus. Interestingly,
this result is not well-known even in the distributed computing community,
and we believe that the SDN paradigm offers a very practical
application for it. We believe that our approach nicely complements
existing, more theoretical literature on distributed control planes; for example,
it shows how to realize STN~\cite{stn}.

\subsection{Organization}

The remainder of this paper is organized as follows.
In Section~\ref{sec:background}, we provide the theoretical
background on our approach. Section~\ref{sec:realization}
then discusses the basic implementation of this approach in Openflow.
Section~\ref{sec:extensions} discusses practical issues and extensions.
After reviewing related literature in Section~\ref{sec:relwork},
we conclude our work in
Section~\ref{sec:conclusion}.

\section{Fundamental Idea}\label{sec:background}

idea: multi-write consensus, a not well-known result!

\section{One-Touch Policy Update}\label{sec:realization}

SDN switch configuration space as shared memory

@@

@@

\section{Discussion and Extensions}\label{sec:extension}

\subsection{Dealing with multiple updates}

@@

\section{Related Work}\label{sec:relwork}

ideas: distributed control planes

ideas: known consensus results and impossibilities

ideas: the power of in-band

\section{Conclusion}\label{sec:conclusion}

ideas: the missing link for STN, but also others

{
\bibliographystyle{abbrv}
\bibliography{references}  % main.bib is the name of the Bibliography in this case
}
